---
title: "EDA- Swire Coca Cola"
author: "Sakshi Pandey"
date: "2024-09-05"
output: 
  html_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
```

## 1. Load Libraries and Data

```{r}
# Load necessary libraries
library(tidyverse) # For data manipulation and visualization
library(lubridate) # For handling date and time data

# Load the CSV file (Replace the path with the correct file location)
iwc_data <- read_csv("IWC_Work_Orders_Extract.csv")

# Check the structure of the data
glimpse(iwc_data)

# Summarize basic statistics
summary(iwc_data)

```

## 2. Data Cleaning and Exploration

```{r}
# Check for missing values
missing_values <- colSums(is.na(iwc_data))
missing_values

# Handle missing data (e.g., removing or imputing)
# You can decide to remove rows with many missing values or impute them based on context
iwc_data_clean <- iwc_data %>% drop_na() # This drops rows with any missing values

# Convert date columns to date format for further analysis
iwc_data_clean <- iwc_data_clean %>%
  mutate(
    EXECUTION_START_DATE = ymd(EXECUTION_START_DATE),
    EXECUTION_FINISH_DATE = ymd(EXECUTION_FINISH_DATE),
    ACTUAL_START_TIME = hms(ACTUAL_START_TIME),
    ACTUAL_FINISH_TIME = hms(ACTUAL_FINISH_TIME)
  )

# Basic visualizations to understand the downtime trends
# Histogram of actual work time
ggplot(iwc_data_clean, aes(x = ACTUAL_WORK_IN_MINUTES)) +
  geom_histogram(binwidth = 50, fill = "blue", color = "white") +
  labs(title = "Distribution of Actual Work Time in Minutes", x = "Work Time (minutes)", y = "Frequency")

# Bar plot of maintenance types
ggplot(iwc_data_clean, aes(x = MAINTENANCE_TYPE_DESCRIPTION)) +
  geom_bar(fill = "green") +
  labs(title = "Count of Maintenance Types", x = "Maintenance Type", y = "Count")

# Time series plot of downtimes by date
ggplot(iwc_data_clean, aes(x = EXECUTION_START_DATE, y = ACTUAL_WORK_IN_MINUTES)) +
  geom_line() +
  labs(title = "Downtimes Over Time", x = "Date", y = "Actual Work Time (minutes)")

```


## 3. Exploratory Data Analysis (EDA)

### 3.1. Step 1: Understanding the Data Structure
First, we will take a closer look at the structure of the data, especially focusing on key variables related to downtime and machine performance

```{r}
# Check the structure again for reference
str(iwc_data_clean)

# Check for unique values in categorical variables
unique(iwc_data_clean$PLANT_ID)
unique(iwc_data_clean$MAINTENANCE_TYPE_DESCRIPTION)
unique(iwc_data_clean$ORDER_DESCRIPTION)

# Summarize downtime by plant
plant_summary <- iwc_data_clean %>%
  group_by(PLANT_ID) %>%
  summarize(
    total_downtime = sum(ACTUAL_WORK_IN_MINUTES, na.rm = TRUE),
    avg_downtime = mean(ACTUAL_WORK_IN_MINUTES, na.rm = TRUE),
    num_orders = n()
  )

# View the summary of downtime per plant
print(plant_summary)


```


### 3.2. Step 2: Visualizing Key Patterns
We'll now visualize some of the key relationships in the data.
The boxplot helps identify how different maintenance types impact the downtime.

```{r}
# Visualize downtime per plant
ggplot(plant_summary, aes(x = PLANT_ID, y = total_downtime)) +
  geom_col(fill = "purple") +
  labs(title = "Total Downtime by Plant", x = "Plant ID", y = "Total Downtime (minutes)")

# Visualize downtime distribution across different maintenance types
ggplot(iwc_data_clean, aes(x = MAINTENANCE_TYPE_DESCRIPTION, y = ACTUAL_WORK_IN_MINUTES)) +
  geom_boxplot(fill = "orange") +
  labs(title = "Downtime Distribution by Maintenance Type", x = "Maintenance Type", y = "Actual Work Time (minutes)")

```


### 3.3. Step 3: Time Series Analysis of Downtime
We can also explore how downtimes vary over time.
This time series plot will help us understand if there are any trends or seasonality in machine downtimes.

```{r}
# Group data by date and summarize downtime
time_series <- iwc_data_clean %>%
  group_by(EXECUTION_START_DATE) %>%
  summarize(total_downtime = sum(ACTUAL_WORK_IN_MINUTES, na.rm = TRUE))

# Plot downtime over time
ggplot(time_series, aes(x = EXECUTION_START_DATE, y = total_downtime)) +
  geom_line(color = "blue") +
  labs(title = "Total Downtime Over Time", x = "Date", y = "Total Downtime (minutes)")

```


### 3.4. Step 4: Correlation Analysis

We will now check for correlations between numerical variables (like actual work time, plant ID, etc.) to see if any relationships stand out.

While plant ID is categorical, R will treat it as a numerical factor in this case, and we can check if any relationships between plant and downtime exist.

```{r}
# Convert PLANT_ID to numeric if needed
iwc_data_clean <- iwc_data_clean %>%
  mutate(PLANT_ID = as.numeric(as.factor(PLANT_ID)))  # Convert PLANT_ID to numeric

# Select numerical columns
numerical_cols <- iwc_data_clean %>%
  select(ACTUAL_WORK_IN_MINUTES, PLANT_ID)

# Check for missing values
missing_values_numerical <- colSums(is.na(numerical_cols))
print("Missing values in numerical columns:")
print(missing_values_numerical)

# Remove rows with missing values
numerical_cols_clean <- na.omit(numerical_cols)

# Compute the correlation matrix
cor_matrix <- cor(numerical_cols_clean)
print("Correlation matrix:")
print(cor_matrix)

```



### 3.5. Step 5: Handling Missing Data
Let's address any missing values in the data and decide how to deal with them (either remove or impute them).

```{r}
# Check again for missing values
missing_values <- colSums(is.na(iwc_data_clean))
print(missing_values)

# We can choose to remove rows with too many missing values or impute them
# For this step, let's remove rows where ACTUAL_WORK_IN_MINUTES is missing, as it's a key variable
iwc_data_clean <- iwc_data_clean %>% filter(!is.na(ACTUAL_WORK_IN_MINUTES))

# Confirm that missing values are handled
sum(is.na(iwc_data_clean$ACTUAL_WORK_IN_MINUTES))

```



### 3.6. Step 6: Results from EDA
Based on the exploration above, here are some key findings:

1. Plant-specific downtime: Some plants experience significantly higher downtime than others, which may be related to their size or operational demands.

2. Maintenance types: Preventive maintenance tends to have less downtime compared to unplanned or corrective maintenance, which supports the idea that a predictive model could reduce unplanned downtimes.

3.Time trends: Downtimes may show seasonal or operational patterns that could be leveraged for prediction.

4. Missing data: Missing values, especially in key columns like downtime, need to be handled carefully (imputation or exclusion)


### 3.7. Step 7: Preparing Data for Modeling
Before proceeding to the modeling phase, we will prepare the dataset by creating additional features, such as:

1. Categorical encoding: Convert categorical variables (e.g., MAINTENANCE_TYPE_DESCRIPTION, PLANT_ID) into numerical values.

2. Time features: Extract features like day of the week, month, etc., from the timestamp columns.

3. Lag features: Create lag variables to see if previous downtimes can help predict future ones.


```{r}
# Convert categorical variables into factors
iwc_data_clean <- iwc_data_clean %>%
  mutate(
    PLANT_ID = as.factor(PLANT_ID),
    MAINTENANCE_TYPE_DESCRIPTION = as.factor(MAINTENANCE_TYPE_DESCRIPTION)
  )

# Create new time-based features
iwc_data_clean <- iwc_data_clean %>%
  mutate(
    weekday = wday(EXECUTION_START_DATE, label = TRUE),
    month = month(EXECUTION_START_DATE, label = TRUE)
  )

# Create lag features (e.g., previous downtime)
iwc_data_clean <- iwc_data_clean %>%
  arrange(EXECUTION_START_DATE) %>%
  group_by(PLANT_ID) %>%
  mutate(lag_downtime = lag(ACTUAL_WORK_IN_MINUTES, 1, order_by = EXECUTION_START_DATE))

# View the updated dataset
head(iwc_data_clean)

```

```{r}
## Export Cleaned Data

# Define the path where the CSV will be saved
csv_path <- "iwc_data_clean.csv"

# Write the cleaned data to a CSV file
write_csv(iwc_data_clean, csv_path)

# Confirm the file has been written
print(paste("CSV file saved to:", csv_path))

```

